---
title: "A Statistical Analysis of Integration and the Negro League"
author: "Paul A. Hodgetts"
date: "09/04/2021"
output: 
  pdf_document:
    latex_engine: lualatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=F, message=F, warning=F}
# uncomment any packages that require installation

# installs tidyverse package for working with data
# version 1.3.1
#install.packages("tidyverse")

# installs here package for accessing directories
# version 1.0.1
#install.packages("here")

# installs showtext package for loading Google fonts
# version 0.9-2
#install.packages("showtext")
```

```{r libraries, echo=F, message=F, warning=F}
# load tidyverse package
# version 1.3.0
library(tidyverse)

# load here package
# version 1.0.1
library(here)

# load showtext package
# version 0.9-2
library(showtext)
```

```{r showtext, message=F, warning=F, echo=F}
# access Jost font from Google fonts
font_add_google(name = "jost", family = "jost-sans-serif")

# load font
showtext_auto()
```

```{r load data, echo=F, message=F, warning=F}
source(here("script/rmarkdown/load_data.R"))
```

```{r prep data, echo=F, warning=F, message=F}
source(here("script/rmarkdown/prep_data.R"))
```

```{r count data, message=F, echo=F, warning=F}
source(here("script/rmarkdown/nl_mlb_count.R"))
```

# Introduction

On April 15th, 1947, Jackie Robinson made his debut for the Brooklyn Dodgers and broke the Major League Baseball (MLB) 'Colour Line' that had been established as an unwritten agreement amongst league owners preventing non-white players from playing in the MLB. Although not the first Black man to play in the major leagues, historians consider William Edward White the first Black player and Moses Fleetwood Walker as the first openly Black player, Robinson was the first to break the established segregation and usher in an era of integration for the MLB. Indeed, Moses Fleetwood Walker was the last Black player in the Majors and it was the refusal of Cap Anson to take the field against Fleetwood Walker that led to the segragation of baseball and establishment of the "Colour Line".

Before joining the Brooklyn Dodgers organization in 1946, Robinson spent the 1945 baseball season playing for the Kansas City Monarchs of the Negro American League. While various iterations of the Negro Leagues had existed, it is generally agreed upon by baseball historians that the period of time that is referred to by use of the Negro Leagues is between the 1920 and the establishment of the Negro National League and the early 1960's with the folding of the Negro American League. In a time during which Black players were not allowed to play in the MLB, the Negro Leagues provided a place for Black players to play and be paid for their efforts. Over the seasons, the Negro Leagues saw such stars as Cool Papa Bell, Satchel Paige, Josh Gibson, Ernie Banks, Willie Mays, and Henry Aaron to name a few. Like Robinson, some of these players saw time in the MLB. However, the integration of the MLB did not eliminate the racism these players experienced, which continued from the fans and other players. Moreover, many players from the Negro Leagues were never given a fair opportunity to play in the MLB, while those who retired from baseball before the 1947 season were never even given a chance. 
